---
title: "UHC Indicator Reference Guide"
subtitle: "Version VERSION_PLACEHOLDER"
date: today
date-format: "DD MMMM YYYY"
format:
  pdf:
    pdf-engine: xelatex
    geometry: "margin=2cm, top=2.5cm"
    toc: true
    toc-depth: 2
    toc-title: "Contents"
    include-in-header:
      text: |
        \usepackage{longtable}
        \usepackage{array}
        \usepackage{colortbl}
        \usepackage{xcolor}
        \usepackage{booktabs}
        \providecommand{\tightlist}{\setlength{\itemsep}{2pt}\setlength{\parskip}{0pt}}
params:
  ids:        ["apx", "tat", "age", "cnc", "dlv", "cur"]
  categories: ["Claim Flow", "Claim Flow", "Claim Flow", "Health Services", "Health Services", "Claim Flow"]
  def_dir:    "definitions"
  version:    "1.0"
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
fields <- list(
  list(key = "definition",      label = "Definition"),
  list(key = "numerator",       label = "Numerator"),
  list(key = "denominator",     label = "Denominator"),
  list(key = "unit_of_measure", label = "Unit of Measure"),
  list(key = "data_source",     label = "Data Source"),
  list(key = "frequency",       label = "Frequency"),
  list(key = "disaggregation",  label = "Disaggregation"),
  list(key = "notes_caveats",   label = "Data Quality Checks")
)

tex_esc <- function(x) {
  if (is.null(x)) return("")
  x <- as.character(x)
  x <- gsub("\\\\", "\\\\textbackslash{}", x)
  x <- gsub("([&%$#_{}])", "\\\\\\1", x)
  x <- gsub("~",   "\\\\textasciitilde{}",  x)
  x <- gsub("\\^", "\\\\textasciicircum{}", x)
  x
}

pandoc_bin <- tryCatch(rmarkdown::pandoc_exec(), error = function(e) Sys.which("pandoc"))
md_to_tex <- function(x) {
  if (is.null(x) || !nzchar(trimws(x))) return("")
  tmp <- tempfile(fileext = ".md")
  writeLines(enc2utf8(as.character(x)), tmp)
  out    <- system2(pandoc_bin, c(shQuote(tmp), "-t", "latex"),
                    stdout = TRUE, stderr = FALSE)
  result <- paste(out, collapse = "\n")
  result <- gsub("\\\\begin\\{longtable\\}(\\[[^]]*\\])?\\{([^}]*)\\}",
                 "\\\\begin{tabular}{\\2}", result, perl = TRUE)
  result <- gsub("\\\\end\\{longtable\\}",  "\\\\end{tabular}", result)
  result <- gsub("\\\\endhead\n?",          "", result)
  result <- gsub("\\\\endfirsthead\n?",     "", result)
  result <- gsub("\\\\endlastfoot\n?",      "", result)
  result <- gsub("\\\\noalign\\{\\}",       "", result)
  result
}

render_indicator_table <- function(id) {
  path   <- file.path(params$def_dir, paste0(id, ".qmd"))
  lines  <- readLines(path, warn = FALSE, encoding = "UTF-8")
  delims <- which(trimws(lines) == "---")
  meta   <- yaml::yaml.load(paste(lines[(delims[1]+1):(delims[2]-1)], collapse = "\n"))

  header <- paste0(
    tex_esc(if (!is.null(meta$indicator_code)) meta$indicator_code else ""),
    if (!is.null(meta$indicator_code) && !is.null(meta$indicator_name)) ": " else "",
    tex_esc(if (!is.null(meta$indicator_name)) meta$indicator_name else id)
  )

  # Section heading for TOC (invisible rule keeps spacing tight)
  cat("\\subsection*{", header, "}\n\\addcontentsline{toc}{subsection}{", header, "}\n\n", sep = "")

  cat(
    "\\begin{longtable}",
    "{|>{\\bfseries\\small\\raggedright\\columncolor[HTML]{EEF2F7}}p{3.4cm}",
    "|>{\\small}p{12cm}|}",
    "\n\\hline\n",
    sep = ""
  )
  cat(
    "\\rowcolor[HTML]{1A4480}",
    "\\multicolumn{2}{|l|}{\\textcolor{white}{\\large\\bfseries ", header, "}}",
    " \\\\[4pt]\n\\hline\n",
    "\\endfirsthead\n",
    "\\multicolumn{2}{c}{\\small\\textit{(continued from previous page)}} \\\\\n\\hline\n",
    "\\endhead\n",
    sep = ""
  )

  for (f in fields) {
    val <- meta[[f$key]]
    if (!is.null(val) && nzchar(trimws(as.character(val)))) {
      cat(tex_esc(f$label), " & ", md_to_tex(val), " \\\\[3pt]\n\\hline\n", sep = "")
    }
  }
  cat("\\end{longtable}\n\n")
}
```

\newpage

\section*{Abbreviations}
\addcontentsline{toc}{section}{Abbreviations}

\begin{tabular}{|>{\bfseries\small}p{3cm}|>{\small}p{11.8cm}|}
\hline
\rowcolor[HTML]{1A4480} \textcolor{white}{\textbf{Abbreviation}} & \textcolor{white}{\textbf{Meaning}} \\
\hline
CR    & Client Registry Number \\[2pt]\hline
DHA   & Digital Health Agency \\[2pt]\hline
ECCIF & Emergency, Chronic and Critical Illness Fund \\[2pt]\hline
ERP   & Enterprise Resource Planning \\[2pt]\hline
POMSF & Post-Operational Medical and Surgical Fund \\[2pt]\hline
SHA   & Social Health Authority \\[2pt]\hline
SHIF  & Social Health Insurance Fund \\[2pt]\hline
TAT   & Turnaround Time \\[2pt]\hline
UHC   & Universal Health Coverage \\[2pt]\hline
\end{tabular}

\newpage

\section*{Acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}

\textit{[Acknowledgements text.]}

\newpage

```{r indicators}
#| output: asis
ids        <- params$ids
categories <- params$categories
unique_cats <- unique(categories)

for (ci in seq_along(unique_cats)) {
  cat_label <- unique_cats[[ci]]
  cat_tex   <- tex_esc(cat_label)
  cat("\\section*{", cat_tex, "}\n\\addcontentsline{toc}{section}{", cat_tex, "}\n\n", sep = "")

  cat_ids <- ids[categories == cat_label]
  for (i in seq_along(cat_ids)) {
    render_indicator_table(cat_ids[[i]])
    if (i < length(cat_ids)) cat("\\newpage\n\n")
  }
  if (ci < length(unique_cats)) cat("\\newpage\n\n")
}
```
